Centralized Programming for Distributed Control Planes

### 2016-01-27 paper/compilation.tex
editting sections 3-5

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{figures/pipeline}
\caption{Compilation pipeline stages for Propane.}
\label{fig:pipeline}
\end{figure}

@@

The high-level language of \S\ref{sec:propane} is just a thin layer on top of a core, regular-expression-based language for describing preference-based path constraints. Regular expressions are an expressive formalism that have been studied extensively for their utility in describing paths through graphs~\cite{bib:todo}, with applications to networks~\cite{bib:todo}. Our regular expression language differs from others by allowing operators to describe preferences between paths.

@@

  \end{minipage}
  %
  ~~
  \vrule
  ~~
  %
  \begin{minipage}[t]{.5\linewidth}
  \hdr{\large Propane Expansions}{}
  \vspace*{-1\baselineskip}
  %
  \[ \begin{array}{rcl}
    \hline
    any           & = & out^* \cdot in^+ \cdot out^* \\
    internal      & = & in^+ \\
    only(X)       & = & any \cap X^* \\
    never(X)      & = & any \cap (!X)^* \\
    through(X)    & = & out^* \cdot in^* \cdot X \cdot in^* \cdot out^* \\
    after(X)      & = & out^* \cdot (X \cap out) \cdot out^* \cdot in^+ \cdot out^* \\
    before(X)     & = & out^* \cdot in^+ \cdot out^* \cdot (X \cap out) \cdot out^* \\
    end(X)        & = & any \cap (\Sigma^* \cdot X) \\
    start(X)      & = & any \cap (X \cdot \Sigma^*) \\
    exit(X)       & = & (out^* \cdot in^* \cdot (X \cap in) \cdot out \cdot out^*) \cup \\
                  &        & (out^* \cdot in^+ \cdot (X \cap out) \cdot out^*) \\
    enter(X)      & = & (out^* \cdot out \cdot (X \cap in) \cdot in^* \cdot out^*) \cup \\
                  &        & (out^* \cdot (X \cap out) \cdot in^+ \cdot out^*) \\
    link(X,Y)     & = & any \cap (\Sigma^* \cdot X \cdot Y \cdot \Sigma^*) \\
    path(\vec{X}) & = & any \cap (\Sigma^* \cdot X_1 \dots X_n \cdot \Sigma^*) \\
    novalley(\vec{X}) & = & any ~ \cap \\
                  &   & !path(X_2,X_1,X_2) ~ \cap \dots \cap \\
                  &   & !path(X_n,X_{n-1},X_n) \\
  \end{array} \]

@@

The syntax for the RIR is shown in Figure~\ref{fig:rir-syntax}. A policy consists of one or more constraints, each of which consists of a test on the type of route, and a corresponding set of preferred regular paths. Regular paths are simply regular expressions where the base characters are abstract locations - either a router or an Autonomous System (AS) number. Special \textit{in} and \textit{out} symbols are used to refer to any internal or external location respectively.
For example, the constraint
$$prefix=74.125.28.0/24 \mapsto (AS200 \cdot in^+) \gg (AS100 \cdot in^+)$$
describes a more-preferred set of paths for traffic announced by a prefixes no less specific than $74.125.28.0/24$, which starts at AS 200, before entering and staying inside the user's network to get to the destination, and a less-preferred set of paths that start at AS 100 and are otherwise the same. The plus operator $in^+$ stands for $in \cdot in^*$ in the usual way. Tests over route types use standard boolean connectives, and can refer to both prefixes and route community values.

\sysname also supports constraints purely on the control-plane behavior of the BGP routing protocol. Thing like prefix aggregation, which should not affect routing behavior, is an important optimization to reduce routing table size and churn in practice. Aggregation, for example from internal locations and external locations, is specified using the same regular syntax as before:
$$agg(128.17.0.0/16, in \rightarrow out)$$
where the expression $in \rightarrow out$ refers to control messages flowing from any internal location to any external location.

@@

The semantics of RIR can be thought of in terms of ranked paths. Each preference-based regular path constraint (of the form $r_1 \gg \dots \gg r_j$) maps to a set of concrete paths in the network that match one of $r_i$. We denote a network path as a string of abstraction locations (routers or external ASes) of the form: $n_1 n_2 \dots n_k$, and we say regular expression $r$ matches path $p$, if $p \in \mathcal{L}(r)$ and $p$ is a topologically-valid path. We denote the length of a path $p$ as $\abs{p}$. A path $p$ will have a rank:
$$(\min_i \set{ p \in \mathcal{L}(r_i) }, \abs{p})$$
where the rank is lexicographically ordered according to (1) the most preferred regular expression matched, and (2) as a tie breaker, the length of the path. A lower rank indicates a \emph{more} preferred path. The desired semantics is to allow traffic to be sent along any of the most preferred paths for each pair of starting and ending locations that appear in some valid specified path.

@@

The next step is to rewrite the high-level constraints such as \textit{enter} and \textit{exit} according to the equivalences listed in Figure~\ref{fig:rir-syntax}. Since preferences can only occur at the outermost level for an RIR expression, the final step is to lift occurences of the preference operator that occur in the regular expression. For example, the regular expression $a \cdot (b \gg c) \cdot d$ is lifted to $(a \cdot b \cdot d) \gg (a \cdot c \cdot d)$ by distributing the preference over the sequence operator. In general, we employ the follwing equivalence:
%
$$x \odot (y_1 \gg \dots \gg y_n) = x \odot y_1 \gg \dots \gg x \odot y_n$$
%
where $\odot$ stands for an arbitrary regular binary operator. Preferences nested under a unary operator, \textit{star} or \textit{negation}, are flagged by the compiler as invalid policies.

>>>>

\begin{figure}[t!]
\centering
\includegraphics[width=.9\columnwidth]{figures/pipeline}
\caption{Compilation pipeline stages for Propane.}
\label{fig:pipeline}
\end{figure}

@@

The high-level language of \S\ref{sec:propane} is just a thin layer on top of a core, regular-expression language for describing preference-based path constraints. Regular expressions are an expressive formalism that have been studied extensively for their utility in describing paths through graphs~\cite{bib:todo}, with applications to networks~\cite{bib:todo}. Our regular expression language differs from others by allowing operators to describe preferences between paths.

@@

  \end{minipage}
  %
  ~~
  \vrule
  ~~
  %
  \begin{minipage}[t]{.5\linewidth}
  \hdr{\large Propane Expansions}{}
  \vspace*{-1\baselineskip}
  %
  \[ \begin{array}{rcl}
    \hline
    any           & = & out^* \cdot in^+ \cdot out^* \\
    internal      & = & in^+ \\
    only(X)       & = & any \cap X^* \\
    never(X)      & = & any \cap (!X)^* \\
    through(X)    & = & out^* \cdot in^* \cdot X \cdot in^* \cdot out^* \\
    later(X)      & = & out^* \cdot (X \cap out) \cdot out^* \cdot in^+ \cdot out^* \\
    before(X)     & = & out^* \cdot in^+ \cdot out^* \cdot (X \cap out) \cdot out^* \\
    end(X)        & = & any \cap (\Sigma^* \cdot X) \\
    start(X)      & = & any \cap (X \cdot \Sigma^*) \\
    exit(X)       & = & (out^* \cdot in^* \cdot (X \cap in) \cdot out \cdot out^*) \cup \\
                  &        & (out^* \cdot in^+ \cdot (X \cap out) \cdot out^*) \\
    enter(X)      & = & (out^* \cdot out \cdot (X \cap in) \cdot in^* \cdot out^*) \cup \\
                  &        & (out^* \cdot (X \cap out) \cdot in^+ \cdot out^*) \\
    link(X,Y)     & = & any \cap (\Sigma^* \cdot X \cdot Y \cdot \Sigma^*) \\
    path(\vec{X}) & = & any \cap (\Sigma^* \cdot X_1 \dots X_n \cdot \Sigma^*) \\
    novalley(\vec{X}) & = & any ~ \cap \\
                  &   & !path(X_2,X_1,X_2) ~ \cap \dots \cap \\
                  &   & !path(X_n,X_{n-1},X_n) \\
  \end{array} \]

@@

The syntax for the RIR is shown in Figure~\ref{fig:rir-syntax}. A policy consists of one or more constraints, each of which consists of a test on the type of route, and a corresponding set of preferred regular paths. Regular paths are simply regular expressions where the base characters are abstract locations - either a router or an Autonomous System (AS) number. Special \textit{in} and \textit{out} symbols are used to refer to any internal or external location respectively.
For example, the constraint
$$prefix=74.3.28.0/24 \mapsto (as200 \cdot in^+) \gg (as100 \cdot in^+)$$
describes a more-preferred set of paths for traffic announced by a prefix no less specific than $74.125.28.0/24$, which starts at AS 200, before entering and staying inside the user's network to get to the destination, and a less-preferred set of paths that start at AS 100 and are otherwise the same. The plus operator $in^+$ stands for $in \cdot in^*$, at least one internal hop. Tests over route types use standard boolean connectives, and can refer to both prefixes and route community values.

\sysname also supports constraints purely on the control-plane behavior of the BGP routing protocol. For example, prefix aggregation, which should not affect routing behavior, is an important optimization to reduce routing table size and churn in practice. Aggregation, for example from internal locations and external locations, is specified using the same regular syntax as before:
$$agg(128.17.0.0/16, in \rightarrow out)$$
where the expression $in \rightarrow out$ refers to control messages flowing from any internal location to any external location.

@@

The semantics of RIR can be thought of in terms of ranked paths. Each preference-based regular path constraint (of the form $r_1 \gg \dots \gg r_j$) maps to a set of concrete paths in the network that match one of $r_i$. We denote a network path as a string of abstract locations (routers or external ASes) of the form: $n_1 n_2 \dots n_k$. A regular expression $r$ matches path $p$, if $p \in \mathcal{L}(r)$ and $p$ is a topologically-valid path. We denote the length of a path $p$ as $\abs{p}$. A path $p$ will have a rank:
$$(\min_i \set{ p \in \mathcal{L}(r_i) }, \abs{p})$$
where the rank is lexicographically ordered according to (1) the most preferred regular expression matched, and (2) as a tie breaker, the length of the path. A lower rank indicates a \emph{more} preferred path. The desired semantics is to allow traffic to be sent along any of the most preferred paths for each pair of starting and ending locations that appear in some valid specified path.

@@

The next step is to rewrite the high-level constraints such as \textit{enter} and \textit{exit} according to the equivalences listed in Figure~\ref{fig:rir-syntax}. Since preferences can only occur at the outermost level for an RIR expression, the final step is to lift occurences of the preference operator in each regular expression. For example, the regular expression $a \cdot (b \gg c) \cdot d$ is lifted to $(a \cdot b \cdot d) \gg (a \cdot c \cdot d)$ by distributing the preference over the sequence operator. In general, we employ the follwing equivalence:
%
$$x \odot (y_1 \gg \dots \gg y_n) = x \odot y_1 \gg \dots \gg x \odot y_n$$
%
where $\odot$ stands for an arbitrary regular binary operator. Preferences nested under a unary operator, \textit{star} or \textit{negation}, are flagged by the compiler as invalid policies.

### 2016-01-27 paper/motivation.tex
editting sections 3-5

Consider the backbone network in Figure~\ref{fig:example1}. It has three neighbors, a customer $Cust$ , a peer $Peer$, and a provider $Prov$. The policy of this network is shown on the right. It prefers the neighbors in a certain order ($P1$) and does not want to act as a transit between Peer and Prov ($P2$). It prefers to exchange traffic with Cust over $R1$ rather than $R2$ because $R1$ is a cheaper ($P3$). To guard against another AS "hijacking" prefixes owned by Cust, it only sends traffic to them if Cust is on the AS path ($P4$). Finally, to guard against Cust accidentally becoming a transit for Prov, it does not use Cust for traffic that will later traverse Prov ($P5$).

@@

We can now see how devising correct configurations for real, larger networks can be a nightmare. Such networks have many neighbors across multiple commercial-relationship classes, differing numbers of links per neighbor, along with several geography-, neighbor- or prefix-based exceptions to the default behavior. Templates help by keeping preference and community values consistent across routers, but operators must still do much of the conceptually difficult work manually.
\sysname lets operator express such policies easily and intuitively. It them automatically generates per-device import and export filters, local preferences, MED attributes, and community tags to ensure that the policy is implemented correctly under all failure scenarios.

>>>>

Consider the backbone network in Figure~\ref{fig:example1}. It has three neighbors, a customer $Cust$ , a peer $Peer$, and a provider $Prov$. The policy of this network is shown on the right. It prefers the neighbors in a certain order ($P1$) and does not want to act as a transit between Peer and Prov ($P2$). It prefers to exchange traffic with Cust over $R1$ rather than $R2$ because $R1$ is cheaper ($P3$). To guard against another AS "hijacking" prefixes owned by Cust, it only sends traffic to them if Cust is on the AS path ($P4$). Finally, to guard against Cust accidentally becoming a transit for Prov, it does not use Cust for traffic that will later traverse Prov ($P5$).

@@

We can now see how devising correct configurations for real, larger networks can be a nightmare. Such networks have many neighbors across multiple commercial-relationship classes, differing numbers of links per neighbor, along with several geography-, neighbor- or prefix-based exceptions to the default behavior. Templates help by keeping preference and community values consistent across routers, but operators must still do much of the conceptually difficult work manually.
\sysname lets operators express such policies easily and intuitively. It then automatically generates per-device import and export filters, local preferences, MED attributes, and community tags to ensure that the policy is implemented correctly under all failure scenarios.

### 2016-01-27 paper/propane.tex
editting sections 3-5

\sysname simplifies network configuration by automatically generating low-level BGP configurations from a high-level specification of routing policy.
%
\ratul{Does this capture the differences correctly?}
Policies for (distributed) control planes differ from data plane policies~\cite{x,y,z} in two ways. First, they must account for all failures at specification time; there is no controller at runtime, the routers must handle failures by themselves in a compliant manner.
%
In \sysname, we enable such specifications through path preferences, with the semantics that a less-preferred path is taken only when a higher-preference path is unavailable.
%
Second, paths may be under-specified (e.g., "prefer customer" does not indicate a concrete path). The \sysname compiler treats such under-specifications as constraints on the set of paths and automatically compute valid sets based on topology.

@@

The operator configures the network with the abstraction that he or she has centralized control over routing and uses a set of high-level constraints to describe the different routes traffic may (or may not) take and their relative preferences.
In general, \sysname specifications are written modularly via a series
of user declarations.
%These definitions allow users to express the three
%major elements of any \sysname specification:  \emph{prefixes},
%\emph{paths} and \emph{policies}.
For example, to begin specification of the backbone
network, we first express the idea that for routes leading to our customer,
we prefer using $R_1$ over $R_2$ (Policy P3 from Figure~\ref{fig:example1}):
%\begin{lstlisting}[mathescape=true]
%define ExitCust = $exit(R_1 \gg R_2)$
%\end{lstlisting}
\begin{code}
\Define ExitCust = \Exit(R1 \Prefer R2)
\end{code}
The  statement above defines a set of \emph{ranked paths}, which includes
all paths (and only those paths) that exit
our network
through either router \CD{R1} or \CD{R2}.  The paths that exit through \CD{R1}
are preferred (\Prefer) to the paths that exit through $R_2$.

@@

\dpw{Note that I did not use an asterix above but used disjunction.}
The first line states that there is no restriction (\Any)
on how traffic must
traverse the network for global prefixes, aside from the default restriction
that traffic must not pass through the user's network and then loop
back on itself. This means traffic for
\CD{PG*} may be sent either from other routers in the datacenter, or
from external ASes. The second line ensures that traffic for local
prefixes never enters the network from an outside location. This constraint
guarantees that the services remain reachable only to locations
internal to the data center.

As in the backbone example, we can combine these constraints
logically to specify the network-wide policy.
However, in addition to constraints on the shape of paths,
\sysname allows the operator to specify constraints on the BGP control plane.
For instance, a constraint on aggregation is included to ensure that
aggregation for the prefixes in the set \CD{AGG} is performed from 
locations inside the network
to locations outside the network. \todo{Say more about aggregation here.
What prefixes are aggregated in to what blocks?  How is this defined?}

\begin{code}
Ownership \AND{} Routing \AND{} \Agg(AGG, \In \Link \Out)
\end{code}

As before, once \sysname compiles the policy, it is guaranteed to hold under
all possible failure scenarios. In addition, the system can check for
aggregate-induced black holes in the presence of up to $k$ failures.
\todo{Note:  Another comment about failure checking for black holes, which may
not be implemented yet.}

>>>>

Policies for (distributed) control planes differ from data plane policies~\cite{x,y,z} in two ways. First, they must account for all failures at specification time; there is no controller at runtime, the routers must handle failures by themselves in a compliant manner.
%
In \sysname, we enable such specifications through path preferences, with the semantics that a less-preferred path is taken only when a higher-preference path is unavailable in the network.
%
Second, paths may be under-specified (e.g., "prefer customer" does not indicate a concrete path). The \sysname compiler treats such under-specifications as constraints on the set of paths and automatically computes valid sets based on topology.

@@

\sysname allows the operator to configure the network with the abstraction that he or she has centralized control over routing and uses a set of high-level constraints to describe the different routes traffic may (or may not) take and their relative preferences.
In general, \sysname specifications are written modularly via a series
of user declarations.
%These definitions allow users to express the three
%major elements of any \sysname specification:  \emph{prefixes},
%\emph{paths} and \emph{policies}.
For example, to begin specification of the backbone
network, we first express the idea that for routes leading to our customer,
we prefer using $R_1$ over $R_2$ (Policy P3 from Figure~\ref{fig:example1}):
%\begin{lstlisting}[mathescape=true]
%define ExitCust = $exit(R_1 \gg R_2)$
%\end{lstlisting}
\begin{code}
\Define ExitCust = \Exit(R1 \Prefer R2)
\end{code}
The  statement above defines a set of \emph{ranked paths}, which includes
all paths (and only those paths) for which traffic exits our network
through either router \CD{R1} or \CD{R2}.  The paths that exit through \CD{R1}
are preferred (\Prefer) to the paths that exit through $R_2$.

@@

The first line states that there is no restriction (\Any)
on how traffic must
traverse the network for global prefixes, aside from the default restriction
that traffic must not pass through the user's network and then loop
back on itself. This means traffic for
\CD{PG*} may be sent either from other routers in the datacenter, or
from external ASes. The second line ensures that traffic for local
prefixes never enters the network from an outside location. This constraint
guarantees that the services remain reachable only to locations
internal to the data center.

As in the backbone example, we can combine these constraints
logically to specify the network-wide policy.
However, in addition to constraints on the shape of paths,
\sysname allows the operator to specify constraints on the BGP control plane.
For instance, a constraint on aggregation is included to ensure that
aggregation for global prefixes (with aggregate \CD{PG}) is performed from 
locations inside the network to locations outside the network. In this case, both \CD{PG1} and \CD{PG2} will use the aggregate \CD{PG} when advertised outside the datacenter.
\todo{Say more about aggregation here.
What prefixes are aggregated in to what blocks?  How is this defined?}

\begin{code}
Ownership \AND{} Routing \AND{} \Agg(P_G, \In \Link \Out)
\end{code}

As before, once \sysname compiles the policy, it is guaranteed to hold under
all possible failure scenarios. \sysname also checks for failure safety against aggregation-induced black holes due to the presence of aggregation constraints. It can efficiently determine a lower bound on the number of failures necessary to black hole traffic given the user's policy, topology, and placement of specific aggregates.
